---
title: "Practicum II"
output: html_notebook
author: Jingjing Liao, Mao Zhang
---

```{r}
library(RSQLite)
dbcon <- dbConnect(RSQLite::SQLite(), "article.db")
```

Create a normalized relational schema: Articles, Authors, Affiliation, Suffix, Collective, Journals, JournalIssues, PubDates, Season

--Drop the tables if exists--
```{sql connection=dbcon}
DROP TABLE IF EXISTS Articles;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Authors;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Suffix;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Collective;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Journals;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS JournalIssues;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS PubDates;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Season;
```

```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS Articles (
  ArticleId INTEGER PRIMARY KEY NOT NULL,
  JournalId INTEGER,
  Language TEXT,
  ArticleTitle TEXT,
  FOREIGN KEY (JournalId) REFERENCES Journals(JournalId)
)
```

```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS Authors (
  AuthorId INTEGER PRIMARY KEY NOT NULL,
  ArticleId INTEGER,
  LastName TEXT,
  ForeName TEXT,
  Initials TEXT,
  Affiliation TEXT,
  SuffixId INTEGER,
  CollectiveId INTEGER,
  FOREIGN KEY (ArticleId) REFERENCES Articles(ArticleId),
  FOREIGN KEY (SuffixId) REFERENCES Suffix(SuffixId),
  FOREIGN KEY (CollectiveId) REFERENCES Collective(CollectiveId)
)
```

```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS Journals (
  JournalId INTEGER PRIMARY KEY NOT NULL,
  ISSN TEXT,
  JournalIssueId INTEGER,
  Title TEXT,
  ISOAbbreviation TEXT,
  FOREIGN KEY (JournalIssueId) REFERENCES JournalIssues(JournalIssueId)
)
```

```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS JournalIssues (
  JournalIssueId INTEGER PRIMARY KEY NOT NULL,
  Volume TEXT,
  Issue TEXT,
  PubDateId INTEGER,
  FOREIGN KEY (PubDateId) REFERENCES PubDates(PubDateId)
)
```

```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS PubDates (
  PubDateId INTEGER PRIMARY KEY NOT NULL,
  Year TEXT,
  Month TEXT,
  Day TEXT,
  MedlineDate TEXT,
  SeasonId INTEGER,
  FOREIGN KEY (SeasonId) REFERENCES Season(SeasonId)
)
```

xmlfn <- "pubmed-tfm-xml/pubmed22n0001-tf.xml"
xmlfn <- "pubmed-tfm-xml/sampleData.xml"
```{r}
library(XML)
xmlfn <- "pubmed-tfm-xml/pubmed22n0001-tf.xml"
xmlDOM <- xmlParse(xmlfn, validate=T)
root <- xmlRoot(xmlDOM)
root
```

Create data frame for Articles, Authors, Suffix, Collective, Journals, JournalIssues, PubDates, Season

```{r}
Articles.df <- data.frame (
                            ArticleId = integer(),
                            JournalId = integer(),
                            Language = character(),
                            ArticleTitle = character(),
                            stringsAsFactors = F
                          )


Authors.df <- data.frame (
                          AuthorId = integer(),
                          ArticleId = integer(),
                          LastName = character(),
                          ForeName = character(),
                          Initials = character(),
                          Affiliation = character(),
                          SuffixId = integer(),
                          CollectiveId = integer(),
                          stringsAsFactors = F
                        )

Suffix.df <- data.frame (
                          SuffixId = integer(),
                          SuffixName = character(),
                          stringsAsFactors = F
                        )

Collective.df <- data.frame (
                          CollectiveId = integer(),
                          CollectiveName = character(),
                          stringsAsFactors = F
                        )

Journals.df <- data.frame (
                          JournalsId = integer(),
                          ISSN = character(),
                          JournalIssueId = integer(),
                          Title = character(),
                          ISOAbbreviation = character(),
                          stringsAsFactors = F
                        )

JournalIssues.df <- data.frame (
                          JournalIssueId = integer(),
                          Volume = character(),
                          Issue = character(),
                          PubDateId = integer(),
                          stringsAsFactors = F
                        )

PubDates.df <- data.frame (
                          PubDateId = integer(),
                          Year = character(),
                          Month = character(),
                          Day = character(),
                          MedlineDate = character(),
                          SeasonId = integer(),
                          stringsAsFactors = F
                        )

Season.df <- data.frame (
                          SeasonId = integer(),
                          SeasonName = character(),
                          stringsAsFactors = F
                        )

```

Checks if it already exists in the passed data frame. Returns a key to the item if it exists, 0 otherwise
```{r}
elementExists <- function(element, df) {
  n <- nrow(df)
  if (n == 0) {
    return (0)
  }
  for (i in 1:n) {
    if (element == df[i, 2]) {
      return (i)
    }
  }
  return (0)
}
```


```{r}
library(r2r)
authorsHashmap <- hashmap()
```

Function: authorExists
Checks if it already exists in the passed data frame. Returns a key to the item if it exists, 0 otherwise and add it to the authorsHashmap
```{r}
authorExists <- function(name) {
  pk <- query(authorsHashmap, name)
  if (is.null(pk)) {
    pk.author <- nrow(Authors.df) + 1
    insert(authorsHashmap, name, pk.author)
    return (0)
  } else {
    return (pk)
  }
}
```

Function: parseAuthors
```{r}
parseAuthors <- function(authorList)
{
  newAuthor.df <- data.frame (
                          LastName = character(),
                          ForeName = character(),
                          Initials = character(),
                          Affiliation = character(),
                          SuffixId = integer(),
                          CollectiveId = integer(),
                          stringsAsFactors = F
                        )
  n <- xmlSize(authorList)
  
    # extract each of the <Author> nodes under <AuthorList>
  for (m in 1 : n) {
    anAuthor <- authorList[[m]]
    nSize <- xmlSize(anAuthor) # Count how many elements for an Author
    if (nSize > 0) {
      for (i in 1 : nSize) {
      value <- xmlValue(anAuthor[[i]])
      name <- xmlName(anAuthor[[i]])
      if (name == "LastName") {
        newAuthor.df[m, 1] <- value
      } else if (name == "ForeName") {
        newAuthor.df[m, 2] <- value
      } else if (name == "Initials") {
        newAuthor.df[m, 3] <- value
      } else if (name == "AffiliationInfo") {
        affiliation <- xmlValue(anAuthor[[i]][[1]])
        newAuthor.df[m, 4] <- affiliation
      } else if (name == "Suffix") {
        exist <- elementExists(value, Suffix.df)
        if (exist == 0) {
          pk.suffix <- nrow(Suffix.df) + 1
          Suffix.df[pk.suffix, 1] <<- pk.suffix
          Suffix.df[pk.suffix, 2] <<- value
          newAuthor.df[m, 5] <- pk.suffix
        } else {
          newAuthor.df[m, 5] <- exist
        }
        
      } else if (name == "CollectiveName") {
        exist <- elementExists(value, Collective.df)
        if (exist == 0) {
          pk.collective <- nrow(Collective.df) + 1
          Collective.df[pk.collective, 1] <<- pk.collective
          Collective.df[pk.collective, 2] <<- value
          newAuthor.df[m, 6] <- pk.collective
        } else {
          newAuthor.df[m, 6] <- exist
        }
        
      }
      
    }
    }
  
  
  }
  
  return (newAuthor.df)
}
```

```{r}
numArticles <- xmlSize(root)
for (m in 1 : numArticles) {
  article <- root[[m]]
  authors <- article[[1]][[4]]
  author <- parseAuthors(authors)
  nAuthor <- nrow(author)
  if (nAuthor > 0) {
      for (i in 1 : nAuthor) {
        # check if each author already exists
         lastName <- author[i,]$LastName
         foreName <- author[i,]$ForeName
         name <- paste(lastName, " ", foreName)
         pk.author <- authorExists(name)
        if (pk.author == 0) {
        # does not exist, so add
        pk.author <- nrow(Authors.df) + 1
        Authors.df[pk.author,3:ncol(Authors.df)] <- author[i,]
        Authors.df[pk.author,1] <- pk.author
       }
     }
  }

}

```

```{r}
article <- root[[1]]
authors <- article[[1]][[4]]
author <- parseAuthors(authors)
author
```

```{r}
nAuthor <- nrow(author)
lastName <- author[1,]$LastName
firstName <- author[1,]$ForeName
name <- paste(lastName, " ", firstName)
name
```

```{r}
Authors.df
```

```{r}
Suffix.df
```

```{r}
Collective.df
```


```{r}
dbDisconnect(dbcon)
```

