---
title: "R Notebook"
output: html_notebook
---

xmlfn <- "pubmed-tfm-xml/pubmed22n0001-tf.xml"
xmlfn <- "pubmed-tfm-xml/sampleData.xml"

```{r}
library(XML)
library(r2r)
```

```{r}
xmlfn <- "pubmed-tfm-xml/sampleData.xml"
xmlDOM <- xmlParse(xmlfn, validate=T)
root <- xmlRoot(xmlDOM)
```

```{r}
Journals.df <- data.frame (
                          JournalsId = integer(),
                          ISSN = character(),
                          JournalIssueId = integer(),
                          Title = character(),
                          ISOAbbreviation = character(),
                          stringsAsFactors = F
                        )

JournalIssues.df <- data.frame (
                          JournalIssueId = integer(),
                          Volume = character(),
                          Issue = character(),
                          PubDateId = integer(),
                          stringsAsFactors = F
                        )

PubDates.df <- data.frame (
                          PubDateId = integer(),
                          Year = character(),
                          Month = character(),
                          Day = character(),
                          MedlineDate = character(),
                          SeasonId = integer(),
                          stringsAsFactors = F
                        )

Season.df <- data.frame (
                          SeasonId = integer(),
                          SeasonName = character(),
                          stringsAsFactors = F
                        )

```

Checks if it already exists in the passed data frame. Returns a key to the item if it exists, 0 otherwise
```{r}
elementExists <- function(element, df) {
  n <- nrow(df)
  if (n == 0) {
    return (0)
  }
  for (i in 1:n) {
    if (element == df[i, 2]) {
      return (i)
    }
  }
  return (0)
}
```

Function: dataExists
Checks if it already exists in the passed data frame. Returns a key to the item if it exists, 0 otherwise and add it to the hashmap
if an author has the same lastname and forename, we treat it as the same author
if a journal has the same ISSN, and Title, we treat it as the same journal
if a pubDate has the same year, month, day and seasonId, we treat it as the same pubDate
```{r}
dataExists <- function(hashmap, df, data) {
  pk <- query(hashmap, data)
  if (is.null(pk)) {
    pk <- nrow(df) + 1
    insert(hashmap, data, pk)
    return (0)
  } else {
    return (pk)
  }
}
```

```{r}
pubDateHashmap <- hashmap()
```

Function:parsePubDate
```{r}
parsePubDate <- function(pubDate) {
  newPubDate.df <- data.frame (
                          Year = character(),
                          Month = character(),
                          Day = character(),
                          MedlineDate = character(),
                          SeasonId = integer(),
                          stringsAsFactors = F
                        )
  
  nSize <- xmlSize(pubDate)
  for (m in 1 : nSize) {
    value <- xmlValue(pubDate[[m]])
    name <- xmlName(pubDate[[m]])
    if (name == "Year") {
      newPubDate.df[1, 1] <- value
    } else if (name == "Month") {
      newPubDate.df[1, 2] <- value
    } else if (name == "Day") {
      newPubDate.df[1, 3] <- value
    } else if (name == "MedlineDate") {
      newPubDate.df[1, 4] <- value
    } else if (name == "Season") {
      # If season has already exists in Season.df, then return the pk, otherwise, add season to the data frame
      exist <- elementExists(value, Season.df)
      if (exist == 0) {
        pk.Season <- nrow(Season.df) + 1
        Season.df[pk.Season, 2] <<- value
        Season.df[pk.Season, 1] <<- pk.Season
        newPubDate.df[1, 5] <- pk.Season
      } else {
        newPubDate.df[1, 5] <- exist
      }
      
    }
  }
  # Check if the pubDate has already exists in PubDates.df, then return pk (using hashmap to look for value), otherwise add newPubDate.df to PubDates.df and return new pk
  year <- newPubDate.df$Year
  month <- newPubDate.df$Month
  day <- newPubDate.df$Day
  seasonId <- newPubDate.df$SeasonId
  date <- paste(c(year, month, day, seasonId), collapse = " ") 
  pk.date <- dataExists(pubDateHashmap, PubDates.df, date)
  if (pk.date == 0) {
    # does not exist, so add
    pk.date <- nrow(PubDates.df) + 1
    PubDates.df[pk.date,2:ncol(PubDates.df)] <<- newPubDate.df[1,]
    PubDates.df[pk.date,1] <<- pk.date
    
  }
  return (pk.date)
}
```

Function: parseJournalIssue
```{r}
parseJournalIssue <- function(journalIssue) {
  pk.JournalIssue <- nrow(JournalIssues.df) + 1
  volume <- xpathSApply(journalIssue, "./Volume", xmlValue)
  issue <- xpathSApply(journalIssue, "./Issue", xmlValue)
  if (length(volume) == 0) {
    volume <- ""
  }
  if (length(issue) == 0) {
    issue <- ""
  }
  if (xmlName(journalIssue[[1]]) == "PubDate") {
    pubDate <- journalIssue[[1]]
  } else if (xmlName(journalIssue[[2]]) == "PubDate") {
    pubDate <- journalIssue[[2]]
  } else {
    pubDate <- journalIssue[[3]]
  }
  JournalIssues.df[pk.JournalIssue, 1] <<- pk.JournalIssue
  JournalIssues.df[pk.JournalIssue, 2] <<- volume 
  JournalIssues.df[pk.JournalIssue, 3] <<- issue
  JournalIssues.df[pk.JournalIssue, 4] <<- parsePubDate(pubDate)
  return (pk.JournalIssue)
}
```

Function: parseJournals
```{r}
parseJournals <- function(journal) {
  newJournal.df <- data.frame (
                          ISSN = character(),
                          JournalIssueId = integer(),
                          Title = character(),
                          ISOAbbreviation = character(),
                          stringsAsFactors = F
                        )
  
  ISSN <- xpathSApply(journal, "./ISSN", xmlValue)
  if (length(ISSN) == 0) {
    ISSN <- ""
    JournalIssue <- journal[[1]]
  } else {
    JournalIssue <- journal[[2]]
  }
  
  Title <- xpathSApply(journal, "./Title", xmlValue)
  if (length(Title) == 0) {
    Title <- ""
  }
  ISOAbbreviation <- xpathSApply(journal, "./ISOAbbreviation", xmlValue)
  if (length(ISOAbbreviation) == 0) {
    ISOAbbreviation <- ""
  }
  
  newJournal.df[1, 1] <- ISSN
  newJournal.df[1, 2] <- parseJournalIssue(JournalIssue)
  newJournal.df[1, 3] <- Title
  newJournal.df[1, 4] <- ISOAbbreviation
  return (newJournal.df)
}
```

```{r}
journalHashmap <- hashmap()
```

```{r}
numArticles <- xmlSize(root)
for (m in 1 : numArticles) {
  article <- root[[m]]
  Journal <- article[[1]][[1]]
  newJournal <- parseJournals(Journal)
  ISSN <- newJournal$ISSN
  Title <- newJournal$Title
  journalInfo <- paste(c(ISSN, Title), collapse = " ")
  pk.journal <- dataExists(journalHashmap, Journals.df, journalInfo)
  if (pk.journal == 0) {
    # does not exist, so add
    pk.journal <- nrow(Journals.df) + 1
    Journals.df[pk.journal, 2:ncol(Journals.df)] <- newJournal[1,]
    Journals.df[pk.journal, 1] <- pk.journal
  }
}
```




```{r}
Journals.df
```



```{r}
PubDates.df
```
```{r}
Season.df
```

```{r}
JournalIssues.df
```



